# Missing values

We would like to explore some meaningful patterns based on the distribution of missing values ie.NAs from our weather dataset. 
(Note: given a fact that our Bike datasets have no missing values, the weather dataset would be our sole data.table to analyze NAs.)

Data table: dat.dt.RData
Varibles of our concern: 
'min_temp’, ‘max_temp’, ‘max_steady_wind’, ‘total_daily_precipitation’, ‘description’

```{r}
library(data.table)
library(ggplot2)
library(anytime)
load('dat.dt.RData')

dat[,year:=year(date)]
dat[,month:=month(date)]
dat[,weekday:=wday(date)]

vars.considered<-c('min_temp','max_temp','max_steady_wind','total_daily_precipitation','description')

# Functions for heatmaps, which take in matrix
makeheatmap1 <- function(interMAT) {
  reshape2:::melt(interMAT)->tu
  tu$Var2<-ordered(tu$Var2,levels=colnames(interMAT))
  tu$Var1<-ordered(tu$Var1,levels=rev(unique(row.names(interMAT))))
  tu$value<-ordered(tu$value,levels=c(0,1),labels=c('No','Yes'))
  ggplot(tu,aes(x=Var2,y=Var1,fill=value)) + 
    geom_tile(width=0.9,show.legend=T) +
    scale_x_discrete(expand=expansion(0)) +
    scale_y_discrete(expand=expansion(0)) +
    labs(x='',y='',fill='isNA') +
    theme(axis.text.y=element_blank(),axis.text.x=element_text(angle=90,vjust=0.5,hjust=1),axis.ticks.y=element_blank()) +
    scale_fill_manual(values=c('grey','#1569C7'))
}

makeheatmap<-function(interMAT,text=F)
{
       reshape2:::melt(interMAT)->tu
       tu$Var2<-ordered(tu$Var2,levels=colnames(interMAT))
       tu$Var1<-ordered(tu$Var1,levels=rev(unique(row.names(interMAT))))
       
       pp<-ggplot(tu,aes(x=Var2,y=Var1,fill=value)) +
             geom_tile(width=0.9,height=0.96,show.legend=T) +
             scale_x_discrete(expand=expansion(0)) +
             scale_y_discrete(expand=expansion(0)) +
             labs(x='',y='',fill='# NAs') +
             theme(axis.text.x=element_text(angle=90,vjust=0.5,hjust=1))+
             scale_fill_gradient(low='#008080',high='#E41B17')

    if(text) pp<-pp + geom_text(aes(label=value),size=3)
    pp
}
```

## Heatmap for overall pattern

Firstly, we would like to present an overall missing-value distribution of continuous days from 2019-01-01 to 2020-12-31 (total 731 days) and the plot shows:

- X : 6 most relevent variables    Y: missing patterns (each day from 2019-01-01 to 2020-12-31)
- Vertically, variable 'max_steady_wind' has most missing values.
- Horizontally, missing values concentrate at the end of the year 2019 and year 2020

```{r}
what<-dat[,..vars.considered]
inter.rowname<-as.character(dat[,date])

is.na(what)->to
row.names(to)<-inter.rowname
as.integer(to)->to2
matrix(to2,nrow=nrow(to))->to2
dimnames(to2)<-dimnames(to)
makeheatmap1(to2)
```


## Visna() for overall pattern

Visna() is another method to explore the potential patterns of missing-value distribution of continuous days (731 days)

- X : 6 most relevent variables    Y: missing patterns (each day from 2019-01-01 to 2020-12-31)
- Vertically, variable 'max_steady_wind' has the most missing values as it ranks number-one at the bottom column-sort. 
- Horizontally, for missing values patterns, pattern 'max_steady_wind' took place most frequently up to 71 times, which means there are 71 rows of data with 'max_steady_wind' value missing as NA. The next pattern is 'min_temp, max_temp, max_steady_wind, total_daily_precipitation, description' with frequency 5 times, which means 5 rows of data miss values for all variables. The complete pattern-rank goes from top to bottom as the right row-sort shows. 

```{r}
library(data.table)
library(ggplot2)
library(grid)
#library(extracat)

load('dat.dt.RData')

vars.considered<-c('min_temp','max_temp','max_steady_wind','total_daily_precipitation','description')
what<-dat[,..vars.considered]
#visna(what,sort='b',return.data=T,opts=list(mar=c(0.5,0.2)))
```
