---
title: "data transformatin"
author: "Yeqi Zhang"
date: "4/5/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,warning = FALSE,
                      message = FALSE)
chooseCRANmirror(graphics=FALSE, ind=1)
knitr::opts_chunk$set(echo = TRUE)
```
# III.Data Transformation 

## 3.1 Data Acquisitiion 

- Download the weather website by httr and GET function;
- Extract the table data from the website by html_table function from rvest package.
```{r}
library(httr)
library(rvest)
from.url<-'https://i-weather.com/weather/washington/history/monthly-history/?gid=4140963&station=19064&month=TODOMONTH&year=TODOYEAR&language=english&country=us-united-states'
todo<-expand.grid(2019:2020,1:12)
names(todo)<-c('year','month')
todo<-todo[order(todo$year,todo$month),]

apply (todo,1,function(xxx) {
   cat(paste(xxx,collapse=","),"====\n")
   gsub('TODOYEAR',xxx[1],from.url)->inter.url
   gsub('TODOMONTH',xxx[2],inter.url)->inter.url
   
   while (1) {
      GET(inter.url)->inter.web
      if(length(grep('</html>',content(inter.web,'text')))) break;
   }
   
   read_html(inter.web)->inter.webcontent
   html_table(inter.webcontent)->inter.tables
   sapply(inter.tables,function(ttt) "Daily minimum temperature" %in% names(ttt))->table.wanted
   
   inter.tables[[which(table.wanted)]]->table.wanted
   
   return(table.wanted)
}) -> resu

unique(lapply(resu,names))

do.call('rbind',resu)->dat

#library(xls)
#write.xlsx(dat,file='data.xlsx',row.names=F)

```

We download a complete weather dataset with 10 variables of 24 months from year 2019 to 2020, rbind and save them as a data.frame; the data can be exported as an xlsx file or saved as RData format. 


## 3.2 Data Manipulation

- Select 6 most relevent variables out of 10 for later analysis
- Rename variables' column names concicely. For example, change Daily minimum temperature to min_temp.
- A few days of weather data from 2019-2020 are missing, for example, there is no data for 2020-12-30; therefore, we fill in the data of those missing days and mark their values as NAs.
- Save the data in data.table format and order by date. 

```{r}
library(data.table)
library(anytime)

load('dat.RData')
setDT(dat)

dat<-dat[,.(date=anydate(Date),
            min_temp=`Daily minimum temperature`,
            max_temp=`Daily maximum temperature`,
            max_steady_wind=`Maximum steady wind`,
            total_daily_precipitation=`Total daily precipitation`,
            description=Description)]

dat[order(date)]->dat
dat[,min_temp:=as.numeric(gsub('°C','',min_temp))]
dat[,max_temp:=as.numeric(gsub('°C','',max_temp))]
dat[,max_steady_wind:=as.numeric(gsub(' Km/h','',max_steady_wind))]
dat[,total_daily_precipitation:=as.numeric(gsub('mm','',total_daily_precipitation))]
        
dat[,range(date)]

all.days<-seq(anydate("2019-01-01"),anydate('2020-12-31'),by='1 days')
days.miss<-anydate(setdiff(all.days,dat[,date]))  # 5 days missing 

toadd<-data.table(date=anydate(days.miss),min_temp=NA,
                  max_temp=NA,max_steady_wind=NA,total_daily_precipitation=NA,description=NA)

rbind(dat,toadd)->dat

dat[order(date)]->dat
dim(dat)

dat

save(dat,file='dat.dt.RData')
```
